---
title: "Introduction to nobelprizer"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to nobelprizer}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(nobelprizer)
```

# Overview

The `nobelprizer` package provides a simple interface for retrieving
Nobel Prize laureate data from the official Nobel Prize API.

It is designed to support data exploration and visualization, including
analysis of award categories, years, gender, birth countries, age at award,
and prize amounts.

# Get laureate data

The main function is `get_nobel_laureates()`.

```{r example, eval=FALSE}
laureates <- get_nobel_laureates(limit = 10)
laureates
```

The returned data frame contains information such as:

- laureate ID
- name
- gender
- birth date
- death date
- birth country
- affiliation
- award year
- Nobel Prize category
- prize amount
- motivation
- official Nobel Prize profile URL

# Shiny application

The Shiny application can be run from the package source directory with:

```{r shiny, eval=FALSE}
shiny::runApp("inst/app")
```

# Data source

The data are retrieved from the official Nobel Prize API.

For API documentation, see:

<https://www.nobelprize.org/about/developer-zone-2/>
